---
title: 'Manual or automatic: how does affect fuel performance?'
author: "Javier Santibáñez"
date: "Saturday, January 24, 2015"
output: 
  pdf_document:
    fig_caption: yes
---

## Summary

In this report we present the results of a comparison between manual and automatic vehicles, measured in miles per gallon of fuel (mpg). The results shows that automatic cars has a better performance than manual cars. The expected difference was estimated in XXX mpg.

## Introduction

We used a data set that consists of the measurements of `r dim(mtcars)[2]` variables from  observations of cars from `r dim(mtcars)[1]` cars. The variables are the following:

- `mpg`, miles(US)/gallon
- `cyl`, number of cylinders
- `dip`, displacement (cu.in.)
- `hp`,  gross horsepower
- `drat`, rear axie ratio
- `wt`, weigth (lb/1000)
- `qsec` 1/4 mile time
- `vs` V/S
- `am` transmission (automatic, manual)
- `gear`, number of forward gears
- `carb`, number of carburetors

Additionaly, we created a new variable, called `am2`, which is the factor version of `am` with levels *automatic* and *manual*.

In this report we have to address the questions:

1. Is an automatic or manual transmission better for performance?
2. What is the difference in performance between automatic and manual transmissions?

To deal with this questions we used a statistical approach, in specific, we used regression models. 

The first qustion can be easily answered. Figure *Performance vs. transmission* in Appendix suggests that performance, measured in miles per gallon, depends of the car transmissions.

```{r,echo=FALSE}
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(knitr)))
suppressWarnings(suppressMessages(library(corrplot)))
mtcars$am2<-factor(mtcars$am,labels=c("Automatic","Manual"))
```

Despite the previous graphic, we must verify that manual trasmission is better than automatic trasnmission. Also, it is important to consider other variables, because some of them can explain part of the variability. 

Finally, we used the software statistical enviroment `R` and the packages `ggplot2` and `corrplot` for graphics.

## Methodology

The following is a brief description of the analysis performed:

1. Fit a regression model with `mpg` (performance) as output and `am` (transmission) as input.

2. Select other covariables from the data set, considering their correlation with `mpg`.

3. Fit a regression model with `mpg` as output and the set of variables from the preceding step as input.

4. Evaluate and compare models fitted in steps 1 and 4.

## Results

### Step 1. Basic model.
The basic model consists in only one input (`mpg`). We used the next code to fit the model.
```{r,}
fitMod1<-lm(mpg~am2,data=mtcars)
summary(fitMod1)
```

The previous results indicates that there is a highly significant difference in perfomance between automatic and manual transmission, estimated as `r coef(fitMod1)[2]` miles per gallon, on average. With this information we can answer our two questions, but we improve them with the following steps. 

We verify the accuracy of the normal assumption with a *qq-plot*. Figure 2 shows that the residuals are aproximately normal distributed.


### Step 2. Select new variables.

To select new variables to include in the model, we used the correlation matrix via the `corrplot` function. From Figure 3 we can identify a set of variables
highly correlated:

- `r names(mtcars)[2]`
- `r names(mtcars)[3]`
- `r names(mtcars)[4]`
- `r names(mtcars)[5]`

Despite this result, we cannot include the whole set of variables because there are a high correlation between them. Then, we have tho choose one variable from that set. We select the variable `hp` because we can use it as a confounder in a extended model, to reduce variability and to improve the estimation of difference in `mpg` due to transmission. 

### Step 3. Extended model.

Now we extend our basic model adding the variable `hp`. First we fit a model with an interection term with the following code.
```{r}
fitMod2<-lm(mpg~am2+hp+am2*hp,data=mtcars)
summary(fitMod2)
```

The previous results suggest that there is evidence to support the hipotesis that the interaction term is not necessary.

We fit a new model wihtout the interaction term between `am` and `hp`, these are the results:
```{r}
fitMod3<-lm(mpg~am2+hp,data=mtcars)
summary(fitMod3)
```

These results suggest support again that there is a difference in performace between automatic and manual transmission, estimated in `r coef(fitMod3)[2]` miles per gallon, but in the case, we take into account a confounder that explains some of the residual variability.

We evaluated the assuption of normality with a *qq-plot*. Figure 4 shows that there is no problems with normality. Also, to check corralation between residuals an covariables we made a scatter plot. In Figure 5 we showed that the slope of a regression line of residuals over the covariable `hp` is aproximately zero, which suggests that there is no correlation problems.

### Step 4. Comparing models.

We estimated two models. An analysis of variance (ANOVA) shows that the extended model is better than the basic model, with a high level of significance (p-value less than 0.05). 

```{r,results='asis'}
kable(anova(fitMod1,fitMod3))
```

Also, the ratio of the sum of squared residuals of the extended model to the basic model is `r round(sum(fitMod3$residual**2)/sum(fitMod1$residual**2),3)`, which means that the residual variation of the extended model is `r round(sum(fitMod3$residual**2)/sum(fitMod1$residual**2)*100,1)`% smallet than the residual variation of the basic model.

## Conclusions

According to the preceding results we can conclude that:

1. A manual transmission is better than an automatic one, in terms of performance, measured in miles per gallon.

2. There is a difference of `r coef(fitMod3)[2]` miles per gallon between manual and automatic transmissions.



## Appendix

```{r,fig.cap="Performance vs. transmision",echo=FALSE,include=TRUE}
g1<-ggplot(data=mtcars,aes(factor(am2),mpg))
g1+geom_boxplot(aes(fill=am2))+
  xlab("Transmission")+ylab("Miles per gallon")+labs(fill="Transmission")
```

```{r,fig.cap="QQ plot for residuals of basic model",include=TRUE,echo=FALSE}
qqnorm(fitMod1$residual,pch=20,col="steelblue",main="")
```

```{r,fig.cap="Correlation matrix",include=TRUE,echo=FALSE}
cor_mtcars<-cor(mtcars[,-12])
corrplot(cor_mtcars,diag=F)
```

```{r,fig.cap="QQ plot for residuals of extended model",include=TRUE,echo=FALSE}
qqnorm(fitMod3$residual,pch=20,col="steelblue",main="")
```

```{r,fig.cap="Scatter plot: residuals versus gross horse power",include=TRUE,echo=FALSE}
data_plot2<-data.frame(hp=mtcars$hp,Residuals=fitMod3$residual)
g2<-ggplot(data_plot2,aes(x=hp,y=Residuals))
g2+geom_point(color="steelblue",size=3)+geom_smooth(method="lm")+
  xlab("Gross horse power")+ylab("Residual")
```




