---
title: 'Manual or automatic: how does affect fuel performance?'
author: "Javier Santibáñez"
date: "Saturday, January 24, 2015"
output: pdf_document
---

## Summary

In this report we present the results of a comparison between manual and automatic vehicles, measured in miles per gallon of fuel (mpg). The results shows that automatic cars has a better performance than manual cars. The expected difference was estimated in XXX mpg.

## Introduction

We used a data set that consists of the measurements of `r dim(mtcars)[2]` variables from  observations of cars from `r dim(mtcars)[1]` cars.

## Methodology

We use the software statistical enviroment `R` and the package `ggplot2` for graphics.

```{r,echo=FALSE}
suppressWarnings(suppressMessages(library(ggplot2)))
mtcars$am2<-factor(mtcars$am,labels=c("Automatic","Manual"))
```
## Results

### Exploratory analysis
First we estimate asimple mode

```{r}
g<-ggplot(data=mtcars,aes(factor(am2),mpg))
g+geom_boxplot(aes(fill=am2))+
  xlab("Transmission")+ylab("Miles per gallon")+labs(fill="Transmission")+
  ggtitle("Performance vs. transmision")
```

Second, we introduced a 
```{r}
g<-ggplot(data=mtcars,aes(x=mpg,y=hp))+geom_point(color="steelblue",size=5)
g+facet_grid(.~am2)+geom_smooth(method="lm")+
  xlab("Weight (in tons)")+ylab("Miles per gallon")+ggtitle("Performance vs. weigth by transmision")
```
### Inferential analysis
```{r}
fitMod1<-lm(mpg~am2,data=mtcars)
fitMod2<-lm()
```







